The Adaptive Multi-scale Simulation Infrastructure - Design and Application

### 2016-01-11 doc/articles/amsi_biotissue/paper.tex
yet again, more paper work

\usepackage{graphicx}
\usepackage{listings}
\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

@@

Declaration of a coupling relation between two scales is required if any communication between the scales is to occur during the simulation (through the amsi libraries), as the coupling declaration during simulation initialization is used to instantiate data structures specific to scale-coupling operations. This limits memory overhead by not allocating space for unneeded coupling operations.

Scales are declared to be related if some quantities of interest (such as physical tensor field values) are transmitted between them for scale-coupling. This coupling relationship is assumed to be asymmetric, thus if scale A is related to scale B, scale B is not implicitly related to scale A, though this may be separately declared.

A simple configuration file can be seen in listing \ref{amsi_config_file} which declares the existence of two scales, macro and micro\_fo, and tells the amsi intialization procedures to assigne them process sets containig 64 and 8128 processes, respectively (for a total of $2^{13}$ processes). The file also specifies the scale-coupling relations, with macro coupled to micro and micro coupled to macro. Note that the order of specification is not relevant to the operation of amsi.
%
\begin{lstlisting}[caption={amsi\_config file},label=amsi_config_file]
  macro 64
  micro_fo 8128
  #
  macro micro_fo
  micro_fo macro_fo
\end{lstlisting}
%
Typically creation of this file will be automated and added to the shell script used to invoke the simulation, so ensure that process counts passed in throught the configuration file are valid. Alternative specification mechanisms such as ratio specification (equivalently percentage specification) are being considered for future use to prevent the possbility of misallocation of processes (the user not specifying assignment for all processes available to the simulation).

The assignment of processes to a process set is currently static, and processes cannot be reassigned during a simulation. Additionally the assignment of a scale to a process set is static as well. However, scale-tasks are capable of being dynamically created and destroyed during runtime. Additionally, scale-tasks associated with specific subsets of a process set (either a single process or several processes) may be migrated in part or in whole. Migrating a scale-task in part amounts to using load balancing facilities on the associated processes to redistribute all data scale-task from a scale-task to a subset of the previously associated processes, allowing the freed processes to be assigned data from other scale-tasks. Migrating a scale task in whole constitutes using load balancing operations on the process set associated with the entire scale to move all pertinent data between subset of the managed process set. In the current implementation these operations are both collective on the process set associated with the managed scale.

While the primary scale (the scale associated with producing the solution for the overall simulation) is often thought of as controlling the tertiary scales in terms of dynamic instantiation, this is not assumed in the amsi implementation. Any scale is capable of informing the process set associated with a coupled scale of the addition or deletion of individual scale-tasks. At present it is dependent on the developer to implement the capability of a scale to accept additional scale-tasks, though algorithms and hooks for developer-designed distribution algorithms are supplied to ease the implementation.

@@

All communication and planning operations in the amsi libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer. At present an application developer must implement up-/down-scaling operations required for their particular multi-scale use case, hooks will be provided in a future release to automate application of such operations during scale-coupling. The current implementation includes on blocking versions of the scale-coupling communication operations. Asynchronous versions have been designed but are not currently being implemented due to lack of immediate need for the primary test amsi use case.

>>>>

\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

@@

A scale coupling is a relation denoting that scale-coupling communication will take place between two scales at some point during the simulation. Declaration of a coupling relation between two scales is required for communicating scales, as the coupling declaration during simulation initialization is used to instantiate data structures specific to scale-coupling operations. This limits memory overhead by not allocating space for unneeded coupling operations. This coupling relationship is asymmetric, thus if scale A is related to scale B (ie. A produces terms required for B to operate), scale B is not implicitly related to scale A, though this may be separately declared.

A simple configuration file can be seen in listing \ref{amsi_config_file} which declares the existence of two scales, macro and micro\_fo, and tells the amsi intialization procedures to assign them process sets containig 64 and 8128 processes, respectively (for a total of $2^{13}$ processes). The file also specifies the scale-coupling relations, with macro coupled to micro and micro coupled to macro. Note that the order of specification is not relevant to the operation of amsi.
%
\begin{lstlisting}[caption={amsi\_config file},label=amsi_config_file]
  macro 64
  micro_fo 8128
  #
  macro micro_fo
  micro_fo macro_fo
\end{lstlisting}
%

Dynamically creating the required configuration file as part of a shell script invoking the simulation is straightforward. Alternative specification mechanisms such as ratio specification (equivalently percentage specification) are being considered for future implementation to prevent the possibility of misallocating processes (the user not specifying assignment for all processes available to the simulation, or attempting to overallocate processes).

The assignment of processes to a process set is static, and processes cannot be reassigned during a simulation. Additionally the assignment of a scale to a process set is static as well. The application programmer may provide capabilities for scale-tasks to be dynamically created and destroyed at runtime, amsi provides functions to inform coupled scales of such dynamic modifications. Since scale-tasks are an abstraction the amsi systems cannot provide direct operations on scale-tasks, but provide mechanisms for the application to inform amsi of such operations when they occur.

Additionally, scale-tasks associated with specific subsets of a process set (either a single process or several processes) may be migrated in part or in whole. Migrating a scale-task in part amounts to using load balancing facilities on the associated processes to redistribute all data scale-task from a scale-task to a subset of the previously associated processes, allowing the freed processes to be assigned data from other scale-tasks. Migrating a scale task in whole constitutes using load balancing operations on the process set associated with the entire scale to move all pertinent data between subset of the managed process set. Amsi provides operations to enact load-balancing operations from plans such as those generated by Zoltan algorithms.  In the current implementation these operations are collective on the process set associated with the managed scale.

While the primary scale (the scale associated with producing the solution for the overall simulation) is often thought of as controlling the tertiary scales in terms of dynamic instantiation, this is not assumed in the amsi implementation. Any scale is capable of informing the process set associated with a coupled scale of the addition or deletion of individual scale-tasks. At present it is dependent on the developer to implement the capability of a scale to accept additional scale-tasks.

@@

Scale-linking quantities are dynamically registered with the amsi system using a data distribution defined on a scale by the scale-tasks operating at that scale. Production of a communication pattern is required prior to executing a scale-coupling communication, and the pattern must be reconciled (synchronized across the coupling relationship) prior to being used to guide a coupling communication operation (and when the communication pattern is altered by the scale providing the coupling data). Several simple communication pattern production algorithms are provided by amsi, but these are relatively trivial due to the nature of scale-coupling communications and the general nature of the amsi libraries. Hooks are provided for users to implement algorithms for the construction of communication patterns tailored to the specifics of their problem.

A communication pattern based on the relationship of scale $A$ coupled to scale $B$ is thought of most simply as a matrix $C \in \mathbb{N}^{|A|\times|B|}$ (where $|X|$ is the cardinality of the process set currently assigned to scale X). The implementation of the communication pattern uses a sparse format in order to reduce the memory impact as the simulation scales. The quantity $C_{ij}$ represents the quantity of scale-coupling terms (of arbitrary type) to be sent from process $i$ in the process set assigned to $A$ to process $j$ in the process set assigned to $B$.

All communication and planning operations in the amsi libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer. Usage only of implicit ordering was decided upon in order to reduce the memory impact as the simulation scales, as any meta-data which scales with the simulation will necessarily pose a memory issue once scaling hits some threshold.

At present an application developer must implement up-/down-scaling operations required for their particular multi-scale use case, hooks will be provided in future releases to automate application of such operations during scale-coupling. The current implementation includes blocking versions of the scale-coupling communication operations. Asynchronous versions have been designed but are not currently implemented due to lack of immediate need in the primary amsi use case.
